AiPoker - Implementation of a Poker Agent

### 2014-11-03 report/paper.tex

\begin{document}

%% The ``\maketitle'' command must be the first command after the
%% ``\begin{document}'' command. It prepares and prints the title block.

@@

Expectiminimax is the continuation of the Minimax algorithm that allows for stochastic events in the game tree such as drawing cards in a game of poker. Expectiminimax trees add an additional type of nodes, chance nodes. At chance nodes a child $c$ is chosen with some probability $P(c)$. This means that a chance node for drawing a card from a full deck has $52$ children with a uniform probability distribution of $1/52$. The value for a chance node is the weighted average of all its children.

@@

\subsubsection{Monte Carlo Tree Search}
MCTS is a best-first search strategy. Every node in the tree contain two values, an estimate of the expected value $V(P)$ of the reward $r(P)$ and the number of times the node has been visited $n_{i}$. MCTS only starts with the root node and incrementally builds the tree by repeating the following 4 steps:\\*\\*
\textbf{Selection}
\\*Selection is made by starting from the root node and then recursively select nodes until a leaf node is reached (this does not have to be a leaf of the game tree). To determine which node is the optimal one, the upper confidence bounds formula (eq(\ref{eq:upperconfidence})) is used. \cite{levente}:

\begin{equation} \label{eq:upperconfidence}
	v_{i} + C \cdot \sqrt{\frac{\ln N}{n_{i}}}
\end{equation}
The first term $v_{i}$ is the estimated value of the node and is responsible for exploitation of good nodes, the second term is made up by the total number of visits of the parent node $N$ and the total number of visits of the current node $n_{i}$. The constant $C$ allows tuning of the exploration-exploitation trade-off. 
\\*\\*
\textbf{Expansion}
\\*If the leaf node reached is not a terminating node, which means that the game is not over at the current node. Then create one or more child nodes from the node currently at. This step is illustrated in fig(\ref{fig:expansion}).
\begin{figure}[here]
  \begin{center}
    \includegraphics[scale=0.28]{img/tree1.png}
    \caption{\label{fig:expansion} Illustration of how the green node is added to the tree.}
  \end{center}
\end{figure}
\\*\\*
\textbf{Simulation}
\\*The third step in MCTS is the simulation step. Here a game is simulated from the state in the new node until a result is achieved. The value of the reached result is recorded.
\\*\\*
\textbf{Backpropagation}
\\*The last step is to update the expected value and the selection counter for all nodes along the explored path by backpropagating the recorded result from the simulation step. This step is illustrated in fig(\ref{fig:backprop}).
\begin{figure}[here]
  \begin{center}
    \includegraphics[scale=0.28]{img/tree2.png}
    \caption{\label{fig:backprop} Illustration of how the path to the node is updated with the score from the simulation in the whole tree.}
  \end{center}
\end{figure}
When the back propagation has been done the whole chain starts over again starting with the selection stage\cite{Chaslot}. The use of the upper bounds confidence selection method makes the game tree grow asymmetrically as it focus its search in the more promising parts of the tree.

@@

\subsubsection{Randomness}
All previously mentioned algorithms take cards as input and gives the probability of the cards in different ways to make the best decision with the current cards on hand. Since a good poker player needs to be able to bluff, a random function has been implemented in order to give the agent a certain level of unpredictability. The agent will make a random move approximately 5\% of the time.

@@

\subsection{AI}
The AI-file contains all methods necessary for the AI to be able to make a decision about what move to make. The heart of the AI is the method: findBestMove() which initializes a new instance of the SearchTree class and then runs a simulation from the current game state. The AI-file also contains methods for Hand Strength, Hand Potential and Effective Hand Strength. Bucketing, which has been mentioned earlier is also present and is used as a heuristic in the first round before the flop has been presented on the table.

>>>>

\begin{document}
\pagenumbering{roman}
%% The ``\maketitle'' command must be the first command after the
%% ``\begin{document}'' command. It prepares and prints the title block.

@@

Expectiminimax is the continuation of the Minimax algorithm that allows for stochastic events in the game tree such as drawing cards in a game of poker. Expectiminimax trees add an additional type of nodes, chance nodes. At chance nodes a child $c$ is chosen with some probability $P(c)$. This means that a chance node for drawing a card from a deck of cards has $52$ children with a uniform probability distribution of $1/52$. The value for a chance node is the weighted average of all its children. [REF!]

@@

\subsubsection{Monte Carlo Tree Search}
MCTS is a best-first search strategy. Every node $(N)$ in the tree contains two values, an estimate of the expected value $V(r(N))$ of the reward $r(N)$ and the number of times the node has been visited $n_{i}$. MCTS only starts with the root node and incrementally builds the tree by repeating the following 4 steps:\\*\\*
\textbf{Selection}
\\*Selection is made by starting from the root node and then recursively select nodes until a leaf node is reached (this does not have to be a leaf of the game tree). To determine which node is the optimal one, the upper confidence bounds formula (eq(\ref{eq:upperconfidence})) is used. \cite{levente}:

\begin{equation} \label{eq:upperconfidence}
	v_{i} + C \cdot \sqrt{\frac{\ln N}{n_{i}}}
\end{equation}
The first term $v_{i}$ is the estimated value of the node and is responsible for exploitation of good nodes, the second term is made up by the total number of visits of the parent node $N$ and the total number of visits of the current node $n_{i}$. The constant $C$ allows tuning of the exploration-exploitation trade-off. 
\\*\\*
\textbf{Expansion}
\\*If the leaf node reached is not a terminating node, which means that the game is not over at the current node. Then create one or more child nodes from the node currently at. This step is illustrated in fig(\ref{fig:expansion}).
\begin{figure}[here]
  \begin{center}
    \includegraphics[scale=0.28]{img/tree1.png}
    \caption{\label{fig:expansion} Illustration of how the green node is added to the tree.}
  \end{center}
\end{figure}
\\*\\*
\textbf{Simulation}
\\*The third step in MCTS is the simulation step. Here a game is simulated from the state in the new node the game has ended. The value of the reached result (won amount of money) is recorded.
\\*\\*
\textbf{Backpropagation}
\\*The last step is to update the expected value and the selection counter for all nodes along the explored path by backpropagating the recorded result from the simulation step. This step is illustrated in fig(\ref{fig:backprop}).
\begin{figure}[here]
  \begin{center}
    \includegraphics[scale=0.28]{img/tree2.png}
    \caption{\label{fig:backprop} Illustration of how the path to the node is updated with the score from the simulation in the whole tree.}
  \end{center}
\end{figure}
When the back propagation has been done the whole chain starts over again starting with the selection stage\cite{Chaslot}. The use of the upper bounds confidence selection method makes the game tree grow asymmetrically as it focus its search in the more promising parts of the tree.

@@

\subsubsection{Randomness}
All previously mentioned algorithms take cards as input and gives the probability of winning with the cards on hand. Since a good poker player needs to be able to bluff, a random function has been implemented in order to give the agent a certain level of unpredictability. The agent will make a random move approximately 5\% of the time.

@@

\subsubsection{AI}
The AI-file contains all methods necessary for the AI to be able to make a decision about what move to make. The heart of the AI is the method: findBestMove() which initializes a new instance of the SearchTree class and then runs the whole MCTS algorithm. The AI-file also contains methods for Hand Strength, Hand Potential and Effective Hand Strength. Bucketing, which has been mentioned earlier is also present and is used as a heuristic in the first round before the flop has been presented on the table.
